---
title: "poll"
author: "Sarah G. Gotfredsen"
date: "9/12/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

LIST OF PACKAGES REQUIRED
```{r echo=TRUE, message=FALSE, warning=FALSE}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}    # data wrangling
if (!require("gplots")) {install.packages("gplots"); require("gplots")}   #for generating ballon plot
if (!require("ggstatsplot")) {install.packages("ggstatsplot"); require("ggstatsplot")} #for visual chi square by condition
if (!require(lme4)) {install.packages("lme4"); require(lme4)}  
if (!require(lmerTest)) {install.packages("lmerTest"); require(lmerTest)} 
if (!require(effects)) {install.packages("effects"); require(effects)}              # plotting regresion results
if (!require(ordinal)) {install.packages("ordinal"); require(ordinal)}  
if (!require(plyr)) {install.packages("plyr"); require(plyr)}
if (!require(scales)) {install.packages("scales"); require(scales)}  

#more packages 
library(ggplot2)
library(hrbrthemes)
library(scales) # to access breaks/formatting functions
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(forcats)
library(lubridate)
```

Read in csv
```{r echo=TRUE, message=FALSE}
df <- read.csv("ca_recall_polls_2021.csv")
```

glimpse data 
```{r}
str(df)
glimpse(df)
```

filter for just Newson and order by ID 
```{r}
df_clean <- df %>% filter(df$candidate_name == "Remove" | df$candidate_name == "Keep")
df_clean <- df_clean[order(df_clean$question_id),]

```


Plot by startdate. Remeber to remove SE 
```{r}
# x-axis is super cluttered w. individual dates so show by month instead 
df_clean$startdate2 <- mdy(df_clean$startdate)

p <- ggplot(df_clean, aes(x=startdate2, y=pct, color=candidate_name)) + 
  geom_point(size=2.5) +
  geom_smooth(se = FALSE, aes(group = candidate_name))+
  #geom_hline(yintercept = mean(df_clean$pct), color="blue")+
  theme_ipsum()


#fianlly, add title 

p + labs(title = "Do Californians want to remove Gavin Newsom from office?") + labs(caption = "(based on data from FiveThirtyEight)") + labs(x = "Election day") + labs(y = "%")



```


Let's look at the potential replacement candidates 
```{r}
df_other <- df %>% filter(df$candidate_name != "Remove", df$candidate_name != "Keep")

```

calculate mean pct for each candidate
```{r}
#df_other_mean <- data.frame(colMeans(na.rm=TRUE, df_other_wide[sapply(df_other_wide, is.numeric)]))

mean_pct <- df_other %>%
  group_by(candidate_name) %>%
  summarise_at(vars(pct), list(pct = mean))



```


plot mean pct by descending order. 
(Larry in the lead followed by Kevin)
```{r}
mean_pct  %>%
  ggplot( aes(x = reorder(candidate_name, -pct), y = pct)) +
  geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
  xlab("") +
  coord_flip() +
  theme_ipsum()
```

see individual polls for each candidate 
```{r}
df_other %>%
  ggplot( aes(x=pct, color=candidate_name, fill=candidate_name)) +
  geom_histogram(alpha=0.6, binwidth = 5) +
  scale_fill_viridis(discrete=TRUE) +
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum() +
  theme(
    legend.position="none",
    panel.spacing = unit(0.1, "lines"),
    strip.text.x = element_text(size = 8)
  ) +
  xlab("") +
  ylab("Assigned Probability (%)") +
  facet_wrap(~candidate_name)
```


look just at Larry 
```{r}

larry <- df_other %>% filter(candidate_name == "Larry A. Elder")


ggplot(larry, aes(x=startdate, y=pct)) + 
  geom_point(size=2.5) +
  geom_smooth()+
  theme_ipsum()


```


plot pct per candidate over time. 
note: it'd be better just to compute the average 
```{r warning=FALSE}
df_other%>%
  ggplot( aes(x=startdate, y=pct, group=candidate_name, color=candidate_name)) +
    geom_smooth(method = "lm", formula = y ~ splines::bs(x, 3), se = FALSE)+
   theme_ipsum() 
```

